---
title: "Weightlifting Data Issues"
author: "Justace Clutter"
date: "June 16, 2014"
output: html_document
---

The weightlifting dataset offered by Eduardo Velloso and Wallace Ugulino consists of the $x$, $y$, and $z$ acceleration components and the associated euler angles for five different sensors placed in various locations on the weightlifter body or on the weights themselves.  During the study, six different weightlifters were instructed to perform a set of 10 reppititions of a known exercise in five different ways; one correct and four incorrect methods.  The five different methds for the [Unilateral Dumbbell Bicepts Curl](http://en.wikipedia.org/wiki/Biceps_curl) exercise were:

 A  Correct
 B  Throwing the Elbows to the front
 C  Lifting the dumbbell only halfway
 D  Lowering the dumbbell only halfway
 E  Throwing the hips to the front
 

 
```{r, echo=FALSE}
# Setup the Environment ---------------------------------------------------

library(ggplot2)
library(caret)
library(grid)
library(gtable)

rm(list = ls())
```

 
```{r, echo=FALSE, cache=TRUE}
source('/Users/justaceclutter/Documents/RStudio/PracticalMachineLearning/rbind_max.R')

# Load the Training and the Test Datasets ---------------------------------

PML_Complete = read.csv('/Users/justaceclutter/Documents/RStudio/PracticalMachineLearning/Data/WearableComputing_weight_lifting_exercises_biceps_curl_variations.csv', header = TRUE)
PML_Training = read.csv('/Users/justaceclutter/Documents/RStudio/PracticalMachineLearning/Data/pml-training.csv', header = TRUE)
PML_Testing = read.csv('/Users/justaceclutter/Documents/RStudio/PracticalMachineLearning/Data/pml-testing.csv', header = TRUE)


# Fix the time variable ---------------------------------------------------

time = PML_Training$raw_timestamp_part_1 + (PML_Training$raw_timestamp_part_2 * 1e-6)
time = as.POSIXct(time, origin='1970-01-01', tz="UTC")
PML_Training$time = time
rm(time)

time = PML_Testing$raw_timestamp_part_1 + (PML_Testing$raw_timestamp_part_2 * 1e-6)
time = as.POSIXct(time, origin='1970-01-01', tz="UTC")
PML_Testing$time = time
rm(time)

time = PML_Complete$raw_timestamp_part_1 + (PML_Complete$raw_timestamp_part_2 * 1e-6)
time = as.POSIXct(time, origin='1970-01-01', tz="UTC")
PML_Complete$time = time
rm(time)

# Split out the window rows and the non-window rows -----------------------

# Each window is 1 second in length and the start of a new window contains all the averaged values 

PML_Training_Window = subset(PML_Training, new_window == 'yes')
PML_Training_NotWindow = subset(PML_Training, new_window == 'no')
```

```{r, echo=FALSE}
genOverviewPlot = function(name) {
  # Generate a plot of some of the summary statistics -----------------------
  
  plots = list()
  
  # Sensor can be one of the following: arm, belt, dumbbell, forearm
  sensor = 'dumbbell'
  
  # User can be one of the following: "adelmo"   "carlitos" "charles"  "eurico"   "jeremy"   "pedro"
  user = name
  measures = sprintf('%s_%s', c('roll', 'pitch', 'yaw'), sensor)
  measureLabels = c('Roll', 'Pitch', 'Yaw')
  
  ds = subset(PML_Complete, user_name == user)
  minT = min(ds$time)
  maxT = max(ds$time)
  limits = c(minT, maxT)
  limits_m = mean(limits)
  limits_w = diff(limits)
  limits = limits_m + ((c(-1, 1)*0.5)*(limits_w/0.9))
  
  useTrainPlot = ggplot(subset(PML_Training, user_name == user)) + theme_bw() + 
    geom_point(aes(x = time, y = 1, color = classe), size = 0.5) + 
    labs(x = NULL, y = 'Training') +
    theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
    theme(legend.position = 'none') +
    coord_cartesian(xlim = limits) +
    theme(axis.ticks = element_blank(), axis.text = element_blank()) +
    theme(panel.margin = unit(c(0,0,0,0), units = 'cm'), plot.margin = unit(c(0,0,0,0), units = 'cm')) +
    theme(panel.border = element_blank()) +
    theme(axis.title = element_text(size = 8))

  
  useTestPlot = ggplot(subset(PML_Testing, user_name == user)) + theme_bw() + 
    geom_point(aes(x = time, y = 1), size = 0.5) + 
    labs(x = NULL, y = 'Testing') +
    theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
    theme(legend.position = 'none') +
    coord_cartesian(xlim = limits) +
    theme(axis.ticks = element_blank(), axis.text = element_blank()) +
    theme(panel.margin = unit(c(0,0,0,0), units = 'cm'), plot.margin = unit(c(0,0,0,0), units = 'cm')) +
    theme(panel.border = element_blank()) +
    theme(axis.title = element_text(size = 8))

  
  plots[[length(plots) + 1]] = useTrainPlot
  plots[[length(plots) + 1]] = useTestPlot
  
  for(i in 1:length(measures)) {
    rp = ggplot(ds) + theme_bw() +
      geom_point(aes_string(x = 'time', y = measures[i], color = 'classe'), size = 0.5) +
      coord_cartesian(xlim = limits) +
      theme(plot.margin = unit(c(0,0,0,0), units = 'mm'), panel.margin = unit(c(0,0,0,0), units = 'mm')) +
      scale_color_discrete(name = "Class") +
      theme(axis.text = element_text(size = 8), axis.title = element_text(size = 8))
    
    if(i == length(measures)) {  
      rp = rp + labs(x = 'Time (UTC)', y = measureLabels[i]) + theme(legend.position = 'right', legend.key = element_blank())
      } else {
        rp = rp + labs(x = NULL, y = measureLabels[i]) + theme(legend.position = 'none') + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())
        }
    
    plots[[i+2]] = rp
    }
  
  getLegend<-function(a.gplot){
    tmp <- ggplot_gtable(ggplot_build(a.gplot))
    leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
    legend <- tmp$grobs[[leg]]
    return(legend)}
  
  # Extract the legend
  leg = getLegend(plots[[length(plots)]])
  
  # Remove the legend from the last plot
  plots[[length(plots)]] = plots[[length(plots)]] + theme(legend.position = 'none')
  
  gp <- do.call(rbind_max, plots)
  gp <- gtable_add_cols(gp, widths = sum(leg$widths))
  panels <- gp$layout$t[grep("panel", gp$layout$name)]
  # set the relative panel heights 1/3 for the top two
  gp$heights[panels] <- lapply(c(1,1,3,3,3), unit, "null")
  # set the legend justification to top (it's a gtable embedded in a gtable)
  leg[["grobs"]][[1]][["vp"]] <- viewport(just = c(0.5,1))
  gp <- gtable_add_grob(gp, leg, t = 1, l = ncol(gp))
  
  gp
}


```


# Adelmo

```{r, echo=FALSE, dpi = 300, fig.align='center'}

gp = genOverviewPlot('adelmo')
grid.draw(gp)

```

# Carlitos

```{r, echo=FALSE, dpi = 300, fig.align='center'}

gp = genOverviewPlot('carlitos')
grid.draw(gp)

```

# Charles

```{r, echo=FALSE, dpi = 300, fig.align='center'}

gp = genOverviewPlot('charles')
grid.draw(gp)

```

# Eurico

```{r, echo=FALSE, dpi = 300, fig.align='center'}

gp = genOverviewPlot('eurico')
grid.draw(gp)

```

# Jeremy

```{r, echo=FALSE, dpi = 300, fig.align='center'}

gp = genOverviewPlot('jeremy')
grid.draw(gp)

```

# Pedro

```{r, echo=FALSE, dpi = 300, fig.align='center'}

gp = genOverviewPlot('pedro')
grid.draw(gp)

```
